### 2015-05-29 abstract.tex
corrections

\thispagestyle{plain}% This way, we can have the page numbering in the foot.
\selectlanguage{english}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus in Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}
The aim of this thesis is to study the impact on the performance and ressource usage of the implementation of cryptographic protocols on a SoC platform.
The platform consist of the combination of an ARM core with an Altera FPGA, on which Barco Silex' crypto IP cores are programmed.

@@

\thispagestyle{plain}
\selectlanguage{francais}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}
Le but de ce mémoire est d'étudier l'implémentation de protocol cryptographiques sur une plateforme SoC et leur impact sur les performances et l'utilisation des ressources.
La plateforme consiste en la combinaison d'un cœur ARM avec un FPGA d'Altera sur lequel sont programmées des IP cores de Barco Silex.

@@

Les résultats sont sans appel~: le coût imposé par OpenVPN non seulement diminue les performance (56~\% de perte sans chiffrement ni authentification), mais en plus il n'exploite pas le hardware, donnant des résultats encore plus mauvais quand les opérations sont déchargées sur le périphérique.
En revanche, IPsec combiné au hardware utilise quatre fois moins de CPU que le software pour des performances similaires de transfert de fichier.
Concernant les connections TLS, le hardware est capable de traiter jusqu'à 7.8 fois plus de connexions par minute que le software avec des paramètre de sécurité élevés, et ce en utilisant 19 fois moins de CPU.

>>>>

\thispagestyle{plain}% This way, we can have the page numbering in the foot.
\selectlanguage{english}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus in Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}
The aim of this thesis is to study the impact on the performance and ressource usage of the implementation of cryptographic protocols on a SoC platform.
The platform consists of the combination of an ARM core with an Altera FPGA, on which Barco Silex' crypto IP cores are programmed.

@@

\thispagestyle{plain}
\selectlanguage{francais}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
	\normalsize
	\end{verse}
	\begin{center}
		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
	\end{center}
Le but de ce mémoire est d'étudier l'implémentation de protocoles cryptographiques sur une plateforme SoC et leur impact sur les performances et l'utilisation des ressources.
La plateforme consiste en la combinaison d'un cœur ARM avec un FPGA d'Altera sur lequel sont programmées des IP cores de Barco Silex.

@@

Les résultats sont sans appel~: le coût imposé par OpenVPN non seulement diminue les performance (56~\% de perte sans chiffrement ni authentification), mais en plus il n'exploite pas le hardware, donnant des résultats encore plus mauvais quand les opérations sont déchargées sur le périphérique.
En revanche, IPsec combiné au hardware utilise quatre fois moins de CPU que le software pour des performances similaires de transfert de fichier.
Concernant les connections TLS, le hardware est capable de traiter jusqu'à 5.89 fois plus de connexions par minute que le software avec des paramètre de sécurité élevés, et ce en utilisant 19 fois moins de CPU.

### 2015-05-29 conclusion.tex
corrections

Although the present work presents some promissing results, the implementation can certainly be improved in several ways and some further experiments should be conducted.

\paragraph{Driver improvement}
The driver of the BA411E can be made less ressources hungry by improving the initialisation of the descriptors and their linking, but the gain would not be significant enough to justify the time investment at this point.
% TODO refer to a stack analysis of an ipsec file transfer, showing the most expensive instructions. If I remember correctly, it was something like the DMA transfer and DMA something freeing.
A better alternative would be to avoid descriptors altogether by modifiying the interface with the IP so that it can use the scatterlists directly.
We would then spare a lot of DMA mapping instructions and thus some precious cycles on the software side.

As we already remarked in section \ref{sec:implem-ba411e-driver}, the use cases involving IPsec were conducted using a previous revision of the driver still activelly polling the IP for its results.
A better and cleaner way to proceed is to use interruption routines.
However, the kernel does not support their current implementation and panics upon usage.
If one were to be willing to spend the time replacing the active polling by clean asynchronous interruptions, he should be aware of the overhead imposed by an interruption.
In some cases, when the operation is just a few clock cycle long for the IP, an active polling could still be the better way to go.
A more thorough comparison of the mutual trade-off deserves some investigation, and as a starting point, the packet size could be treated as a branching point between the two solutions.

\paragraph{Registering public key verification with the Linux Crypto API}
The BA414E driver is already capable of offloading a large portion of public key operations to the IP core, but only with very specific libraries at the time being -- OpenSSL in our case.
The next step is to register the very same operations with the Linux Crypto API so they can be used without having to rely on a custom OpenSSL engine.
This feature would require to work very closely to the linux kernel developpement.
Indeed, if the signature verification using public key cryptography has been available in the kernel since 2013, a public key encryption API has only been proposed in late April 2015~\cite{crypto-api-pk-encryption} and is still under request for comments.

\paragraph{Conditional offloading in cryptodev}
%Note this is still under heavy criticism. If the hardware lags so far behind, why the relatively good score in openvpn?
The figure \ref{fig:openssl-speed} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
%TODO include a listing to where it could be a good place to branch on the packet size. I already found it.
%TODO at least add a pseudo-code to illustrate this shit.
Using this breakeven point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise. %openssl-1.0.2a/crypto/engine/eng_cryptodev.c
It would probably not be as trivial and beneficial as it may sound:
\begin{itemize}
	\item the encryption contexts would need to be synchronized between the hardware and the software;
	% \item we would already suffer from some context switches when sending the information to the cryptodev driver; %No, we are still in the openssl engine.
	% \item if we always branch to the fallback functions, the time overhead due to crytodev would lower the performance; %No, we don't leave the user space, the overhead is minimum.
	\item as the breakeven point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
\end{itemize}

Such a conditional offloading would be interresting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeats.

@@

However, one library definitely worth to keep an eye on is mbed TLS, formally known as PolarSSL, recently bought by ARM~\cite{2015-arm-buy-polarssl}.
We can expect the future releases of this library to be more optimized for ARM platforms, and maybe the software footprint and overhead to be reduced.

@@

This thesis studied the implementation of cryptographic protocols on a platform with hardware offloading, using realistic use cases.
The analysis of the results lead to the conclusion that a user-space VPN such as OpenVPN is not a good choice when the objective is to profit from a dedicated physical device.
Instead, if one has all the operations he wants registered in the kernel, he should turn toward IPsec.
OpenVPN took over the market ten years ago when IPsec was not a standard feature of the Linux kernel yet and was quite complex to implement.
Technologic inertia left OpenVPN at the top of the tree, but modern infrastructures should favor IPsec.\newline{}

Security is needed everywhere.
But security is ressource-hungry, and consequently increase the power consumption of the platform.
An ARM core is not meant for comple cryptographic operations.
Combining it with an FPGA like Altera did opens a new field of application for such a SoC platform.
Embedding this SoC into a low-energy platform could add cryptogrpahic capabilities in places it was not possible before.

The strength of this systems also lies into its configurability.
It is not a closed and frozen system, the client has a plaform that can evolve and is reconfigurable according to his needs.\newline{}

>>>>

Although the present work presents some promising results, the implementation can certainly be improved in several ways and some further experiments should be conducted.

\paragraph{Driver improvement}
The driver of the BA411E can be made less resources hungry by improving the initialisation of the descriptors and their linking, but the gain would not be significant enough to justify the time investment at this point.
% TODO refer to a stack analysis of an ipsec file transfer, showing the most expensive instructions. If I remember correctly, it was something like the DMA transfer and DMA something freeing.
A better alternative would be to avoid descriptors altogether by modifying the interface with the IP so that it can use the scatterlists directly.
We would then spare a lot of DMA mapping instructions and thus some precious cycles on the software side.

As we already remarked in section \ref{sec:implem-ba411e-driver}, the use cases involving IPsec were conducted using a previous revision of the driver still actively polling the IP for its results.
A better and cleaner way to proceed is to use interruption routines.
However, the kernel does not support their current implementation and panics upon usage.
If one were to be willing to spend the time replacing the active polling by clean asynchronous interruptions, he should be aware of the overhead imposed by an interruption.
In some cases, when the operation is just a few clock cycle long for the IP, an active polling could still be the better way to go.
A more thorough comparison of the mutual trade-off deserves some investigation, and as a starting point, the packet size could be treated as a branching point between the two solutions.

\paragraph{Registering public key verification with the Linux Crypto API}
The BA414E driver is already capable of offloading a large portion of public key operations to the IP core, but only with very specific libraries at the time being -- OpenSSL in our case.
The next step is to register the very same operations with the Linux Crypto API so they can be used without having to rely on a custom OpenSSL engine.
This feature would require to work very closely to the linux kernel development.
Indeed, if the signature verification using public key cryptography has been available in the kernel since 2013, a public key encryption API has only been proposed in late April 2015~\cite{crypto-api-pk-encryption} and is still under request for comments.

\paragraph{Conditional offloading in cryptodev}
%Note this is still under heavy criticism. If the hardware lags so far behind, why the relatively good score in openvpn?
The figure \ref{fig:openssl-speed} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
%TODO include a listing to where it could be a good place to branch on the packet size. I already found it.
%TODO at least add a pseudo-code to illustrate this shit.
Using this break even point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise. %openssl-1.0.2a/crypto/engine/eng_cryptodev.c
It would probably not be as trivial and beneficial as it may sound:
\begin{itemize}
	\item the encryption contexts would need to be synchronized between the hardware and the software;
	% \item we would already suffer from some context switches when sending the information to the cryptodev driver; %No, we are still in the openssl engine.
	% \item if we always branch to the fallback functions, the time overhead due to crytodev would lower the performance; %No, we don't leave the user space, the overhead is minimum.
	\item as the break even point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
\end{itemize}

Such a conditional offloading would be interesting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeats.

@@

However, one library definitely worth to keep an eye on is mbed TLS, formerly known as PolarSSL, recently bought by ARM~\cite{2015-arm-buy-polarssl}.
We can expect the future releases of this library to be more optimized for ARM platforms, and maybe the software footprint and overhead to be reduced.

@@

This thesis studied the implementation of cryptographic protocols on a platform with hardware offloading, using realistic use cases.
The analysis of the results lead to the conclusion that a user-space VPN such as OpenVPN is not a good choice when the objective is to profit from a dedicated physical device.
Instead, if one has all the operations he wants registered in the kernel, he should turn toward IPsec.
OpenVPN took over the market ten years ago when IPsec was not a standard feature of the Linux kernel yet and was quite complex to implement.
Technological inertia left OpenVPN at the top of the tree, but modern infrastructures should favour IPsec.\newline{}

Security is needed everywhere.
But security is resource-hungry, and consequently increase the power consumption of the platform.
An ARM core is not meant for complex cryptographic operations.
Combining it with an FPGA like Altera did open a new field of application for such a SoC platform.
Embedding this SoC into a low-energy platform could add cryptographic capabilities in places it was not possible before.

The strength of this systems also lies into its reconfigurability.
It is not a closed and frozen system, the client has a platform that can evolve and is reconfigurable according to his needs.\newline{}

### 2015-05-29 implementation.tex
corrections

\section{Software}
In order to implement all the required cryptographic protocol, several softwares are needed.
Their role is summarized in table~\ref{tab:soft-role}.

\begin{table}[ht]
\center
\begin{tabular}{|l|l|} \hline
OpenVPN & SSL/TLS \\ \hline
OpenSSH & SSH \\ \hline
OpenSSL & Cryptographic ciphers implementation \\ \hline
Strongswan & IPsec \\ \hline
\end{tabular}
\caption{Role of the softwares}{Each software implements different protocols. OpenSSL is particular since it implements the cryptographic algorithms themselves and is used by the other applications.}
\label{tab:soft-role}
\end{table}

@@

Since OpenVPN aims at network security at the application level, it needs virtual network interface to read and write the data.
The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.

OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
The problem of encapsulating IP into TCP is that inside IP packets, there already are TCP frames.
The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
Hence, encapsulating an IP packet into a TCP packet produces a redundancy by nesting reliability layers.
UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encaspulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.

@@

\noindent At this point, it is important to note that if the figure~\ref{fig:openvpn-packet-flow} shows a workflow in which the encryption taking place before the fragmentation, a quick look at the source code proves the opposite.
A sample of the said code has been writen down in the listing~\ref{list:openvpn-workflow}.

\lstset{language=c}
\begin{lstlisting}[caption={openvpn compress then encrypt -- sample from \texttt{forward.c}. It clearly shows that the order of operations in the packet workflow is compression, then fragmentation and finally encryption.}, label=list:openvpn-workflow, float]% The 'float' option makes the listing unbreakable.
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

@@

An other way to look at this workflow is to completely isolate the OpenVPN block from both hosts, and exporting it to an independent device such as a router.
The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
All their communications would be routed by the router of their local area network, incomming on a physical \texttt{tun0} interface and outputed to the internet on the \texttt{eth0} interface.
Such implementation are used when an adminstrator wishes to secure a whole local area network without bloating each client with a VPN application.
In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
Whatever the method used, the crucial point is that the VPN tunnel usage is transparent to the application.\newline{}

@@

A previous study highlighted that OpenVPN would divide the throughput by two from raw to tunneled without encryption nor authentication~\cite{Hoekstra2011}.
Such a behaviour would not be surprising, considered all the transfers and context switches the data would have to operate and provoke in an OpenVPN tunnel.

\noindent Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN ealier than 2.3.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.

@@

OpenSSH relies on an external cryptographic library for all its security operations.
Up until the version 6.7 published in october 2014, it had to be compiled against OpenSSL.
However, after the infamous security vulnerability heartbleed in april 2014, the developpers took a step to move towards LibreSSL, a fork of OpenSSL managed by OpenBSD  developers.
Still, there is no official support for any other cryptographic library.

@@

\noindent The first customization is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.

\noindent The second has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, skipping the MAC can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to overcome this limitation, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.

@@

OpenSSL can extend its features by supporting cyrptographic engines.
Those engines offer custom implementation for specific ciphers.
In a sense, a crypto engine is to OpenSSL what the crypto kernel modules are to the Linux Crypto API.

@@

% Talk about the assembly implementation, maybe show the difference in an appendice with the C implementation. Not /that/ interresting.

\subsection{Strongswan}
% TODO talk about IKE and SA management
Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the Linux Crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation difficult.
%Use strongswan 5.3.0
Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its user interface is straightforward.
Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready-to-use VPN.

The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPsec ESP tunnel.
The XFRM (read ``transform") framework is implementing IPsec and handles the incomming and outgoing packets for established VPN connections~\cite{rosen2014}.
Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
Depending on the configuration, XFRM uses the AH or ESP kernel module, which in turn calls the crypto API to encrypt and/or sign the IP packet.

We can also clearly see one of the main advantages of IPsec: it works in the kernel space.
Since it does not require a virtual network interface like OpenVPN, the only transfer between the user/kernel space happens when the former wishes to send a packet on the network, passing it to the later -- or \textit{vice versa} for incomming packets.

\begin{figure}[ht]
\Large
\resizebox{\linewidth}{!}{%
\input{ipsec-transfer}
}
\caption{IPsec user/kernel space workflow, using \texttt{ping} as a test case.}{}
\label{fig:ipsec-workflow}
\end{figure}

@@

\begin{description}
	\item[\texttt{aes\_arm}] Assembly implementation of AES. This version is optimized to use the ARMv7 NEON instruction set.
	\item[\texttt{sha256-generic}] C implementation of SHA-256.
	\item[\texttt{gfmult128}] Multiplication in $GF(2^{128})$, needed by the GCM mode.
	\item[\texttt{ghash}] GHASH function needed by the authentication of the GCM mode.
	\item[\texttt{seqiv}] Sequence IV number generator, needed by the CTR and GCM modes.
	\item[\texttt{cbc}] C based CBC mode.
	\item[\texttt{gcm}] C based GCM mode.
	\item[\texttt{crypto\_null}] Null cipher. This basically does nothing on the plaintext, but it is still needed to make the interface available in the kernel.
	\item[\texttt{tun}] TUN/TAP network device, needed by OpenVPN.
	\item[\texttt{xfrm\_user}] XFRM operations.
	\item[\texttt{esp4}] IPv4 ESP implementation. The AH counterparty is also available in the \texttt{ah4} module.
	\item[\texttt{ipcomp}] IP compression module. Needed by IPsec if the option is activated in the configuration.
	\item[\texttt{cryptodev}] Creates \texttt{/dev/crypto}, giving access to the crypto API in the user space.
	\item[\texttt{uio}] User-space I/O driver, allowing to access the hardware memory from the user space, needed by our implementation of the BA414E driver.
\end{description}

@@

\subsection{BA411E Driver}\label{sec:implem-ba411e-driver}
This driver takes the rightern path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
By pluging the driver into the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space applications via the cryptodev driver and OpenSSL engine.

@@

Pressed by exhibitions and the complexity of interfacing the functions with the crypto API, Barco Silex' developmement team took the choice to access the hardware from user space.

\noindent The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
The complete work flow thus needs three elements:
\begin{itemize}
	\item A device driver handling the IRQ, initializing the device and creating a simple file system that will be used from the user space.
	\item The UIO kernel module supervising the memory access of the user space.
	\item The custom OpenSSL silex engine that translates the requests from OpenSSL into operations processable by the IP core.
\end{itemize}

>>>>

\section{Software}
In order to implement all the required cryptographic protocol, several software are needed.
Their role is summarized in table~\ref{tab:soft-role}.

\begin{table}[ht]
\center
\begin{tabular}{|l|l|} \hline
OpenVPN & SSL/TLS \\ \hline
OpenSSH & SSH \\ \hline
OpenSSL & Cryptographic ciphers implementation \\ \hline
Strongswan & IPsec \\ \hline
\end{tabular}
\caption{Role of the software}{Each software implements different protocols. OpenSSL is particular since it implements the cryptographic algorithms themselves and is used by the other applications.}
\label{tab:soft-role}
\end{table}

@@

Since OpenVPN aims at network security at the application level, it needs virtual network interface to read and write the data.
The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two clients, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.

OpenVPN as two choices when it wants to send its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
The problem of encapsulating IP into TCP is that inside IP packets, there already are TCP frames.
The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
Hence, encapsulating an IP packet into a TCP packet produces a redundancy by nesting reliability layers.
UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encapsulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.

@@

\noindent At this point, it is important to note that if the figure~\ref{fig:openvpn-packet-flow} shows a work flow in which the encryption taking place before the fragmentation, a quick look at the source code proves the opposite.
A sample of the said code has been written down in the listing~\ref{list:openvpn-workflow}.

\lstset{language=c}
\begin{lstlisting}[caption={openvpn compress then encrypt -- sample from \texttt{forward.c}. It clearly shows that the order of operations in the packet work flow is compression, then fragmentation and finally encryption.}, label=list:openvpn-workflow, float]% The 'float' option makes the listing unbreakable.
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

@@

Another way to look at this workflow is to completely isolate the OpenVPN block from both hosts, and exporting it to an independent device such as a router.
The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
All their communications would be routed by the router of their local area network, incoming on a physical \texttt{tun0} interface and output to the internet on the \texttt{eth0} interface.
Such implementation are used when an administrator wishes to secure a whole local area network without bloating each client with a VPN application.
In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
Whatever the method used, the crucial point is that the VPN tunnel usage is transparent to the application.\newline{}

@@

A previous study highlighted that OpenVPN would divide the throughput by two from raw to tunnelled without encryption nor authentication~\cite{Hoekstra2011}.
Such a behaviour would not be surprising, considered all the transfers and context switches the data would have to operate and provoke in an OpenVPN tunnel.

\noindent Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN earlier than 2.3.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.

@@

OpenSSH relies on an external cryptographic library for all its security operations.
Up until the version 6.7 published in October 2014, it had to be compiled against OpenSSL.
However, after the infamous security vulnerability heartbleed in April 2014, the developers took a step to move towards LibreSSL, a fork of OpenSSL managed by OpenBSD  developers.
Still, there is no official support for any other cryptographic library.

@@

\noindent The first customization is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in October 2014.}.
Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.

\noindent The second has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, skipping the MAC can be interesting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to overcome this limitation, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.

@@

OpenSSL can extend its features by supporting cryptographic engines.
Those engines offer custom implementation for specific ciphers.
In a sense, a crypto engine is to OpenSSL what the crypto kernel modules are to the Linux Crypto API.

@@

% Talk about the assembly implementation, maybe show the difference in an appendice with the C implementation. Not /that/ interesting.

\subsection{Strongswan}
% TODO talk about IKE and SA management
Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the Linux Crypto API for the cryptographic part, and on user space crypto libraries for the connection negotiation.
Another popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation difficult.
%Use strongswan 5.3.0
Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its user interface is straightforward.
Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready-to-use VPN.

The figure~\ref{fig:ipsec-workflow} illustrates the work flow of Alice communicating with Bob via an IPsec ESP tunnel.
The XFRM (read ``transform") framework is implementing IPsec and handles the incoming and outgoing packets for established VPN connections~\cite{rosen2014}.
Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
Depending on the configuration, XFRM uses the AH or ESP kernel module, which in turn calls the crypto API to encrypt and/or sign the IP packet.

We can also clearly see one of the main advantages of IPsec: it works in the kernel space.
Since it does not require a virtual network interface like OpenVPN, the only transfer between the user/kernel space happens when the former wishes to send a packet on the network, passing it to the later -- or \textit{vice versa} for incoming packets.

\begin{figure}[ht]
\Large
\resizebox{\linewidth}{!}{%
\input{ipsec-transfer}
}
\caption{IPsec user/kernel space work flow, using \texttt{ping} as a test case.}{}
\label{fig:ipsec-workflow}
\end{figure}

@@

\begin{description}
	\item[\texttt{aes\_arm}] Assembly implementation of AES. This version is optimized to use the ARMv7 NEON instruction set.
	\item[\texttt{sha256-generic}] C implementation of SHA-256.
	\item[\texttt{gfmult128}] Multiplication in $GF(2^{128})$, needed by the GCM mode.
	\item[\texttt{ghash}] GHASH function needed by the authentication of the GCM mode.
	\item[\texttt{seqiv}] Sequence IV number generator, needed by the CTR and GCM modes.
	\item[\texttt{cbc}] C based CBC mode.
	\item[\texttt{gcm}] C based GCM mode.
	\item[\texttt{crypto\_null}] Null cipher. This basically does nothing on the plaintext, but it is still needed to make the interface available in the kernel.
	\item[\texttt{tun}] TUN/TAP network device, needed by OpenVPN.
	\item[\texttt{xfrm\_user}] XFRM operations.
	\item[\texttt{esp4}] IPv4 ESP implementation. The AH counterparty is also available in the \texttt{ah4} module.
	\item[\texttt{ipcomp}] IP compression module. Needed by IPsec if the option is activated in the configuration.
	\item[\texttt{cryptodev}] Creates \texttt{/dev/crypto}, giving access to the crypto API in the user space.
	\item[\texttt{uio}] User-space I/O driver, allowing accessing the hardware memory from the user space, needed by our implementation of the BA414E driver.
\end{description}

@@

\subsection{BA411E Driver}\label{sec:implem-ba411e-driver}
This driver takes the righten path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
By pluging the driver in the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space applications via the cryptodev driver and OpenSSL engine.

@@

Pressed by exhibitions and the complexity of interfacing the functions with the crypto API, Barco Silex' development team took the choice to access the hardware from user space.

\noindent The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the path to the left of figure~\ref{fig:os-path-generic-b}.
The complete work flow thus needs three elements:
\begin{itemize}
	\item A device driver handling the IRQ, initializing the device and creating a simple file system that will be used from the user space.
	\item The UIO kernel module supervising the memory access of the user space.
	\item The custom OpenSSL silex engine that translates the requests from OpenSSL into operations processable by the IP core.
\end{itemize}

### 2015-05-29 introduction.tex
corrections

\chapter{Introduction}
% What is hardware offloading?
In 1965, Gordon Moore, co-founder of Intel, stated that for the same footprint, the number of transistors in an integrated circuit should double every two years~\cite{moore1965}.
It became the Moore's law, an empirical prediction that drived inovation for fifty years.
Nowadays, it is reaching its limits, as Intel in delaying the manufacturing of its next-generation processors.

@@

\noindent A report of the Natural Resources Defense Council~\cite{nrdc2014} forecasts the US datacenter electrity bill to reach \$13.7 billion by 2020.
As CPUs are largely responsible for the power consumption, it makes their design critical.
A solution is to use a hybrid architecture, combining a general-purpose micro-processor with an integrated circuit configurable on-the-fly.
This is a path Intel is exploring with its new Xeon product line, a processor aimed at servers, that combines a Xeon CPU with an FPGA from Altera\cite{xeon2014}.\newline{}

@@

% Power is critical nowadays, esapcially in datacenters.
% Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
% Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.

@@

\section{Structure}
The present document will begin with the technical ground necessary to this work.
After a presentation on how to drive a hardware device from the operating system and details on device driver design, we will adress the cryptography.
Only most relevant part of cryptography will be detailed: message digests, symmetric cryptography and more specifically the CBC and GCM modes of operation, and asymmetric cryptography with an accent on RSA and Diffie-Hellman.

>>>>

\chapter{Introduction}
% What is hardware offloading?
In 1965, Gordon Moore, co-founder of Intel, stated that for the same footprint, the number of transistors in an integrated circuit should double every two years~\cite{moore1965}.
It became the Moore's law, an empirical prediction that drived innovation for fifty years.
Nowadays, it is reaching its limits, as Intel in delaying the manufacturing of its next-generation processors.

@@

\noindent A report of the Natural Resources Defense Council~\cite{nrdc2014} forecasts the US datacenter electricity bill to reach \$13.7 billion by 2020.
As CPUs are largely responsible for the power consumption, it makes their design critical.
A solution is to use a hybrid architecture, combining a general-purpose micro-processor with an integrated circuit configurable on-the-fly.
This is a path Intel is exploring with its new Xeon product line, a processor aimed at servers, that combines a Xeon CPU with an FPGA from Altera\cite{xeon2014}.\newline{}

@@

% Power is critical nowadays, esapcially in datacenters.
% Several solutions exists: abandonning general purpose processors altogether and turn towards ARM cores, or add an FPGA to the processor for example.
% Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.

@@

\section{Structure}
The present document will begin with the technical ground necessary to this work.
After a presentation on how to drive a hardware device from the operating system and details on device driver design, we will address the cryptography.
Only most relevant part of cryptography will be detailed: message digests, symmetric cryptography and more specifically the CBC and GCM modes of operation, and asymmetric cryptography with an accent on RSA and Diffie-Hellman.
